---
title: "Caterpillar Abundance Analysis"
author: "Edward R. Ivimey-Cook"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: show
---

## Introduction

This report examines how caterpillar abundance varies across habitat types,
using a Poisson generalised linear model and a boxplot of counts by habitat.

### Packages  
```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Load packages  
library(dplyr)
library(ggplot2)
library(here)
```

## Data

### Import

Caterpillar counts and their associated habitat are read in from a CSV file,
using `here::here()` so the path resolves without relying on local file
paths.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Load caterpillar abundance data (w/o local file paths)
data <- read.csv(here("caterpillar_data", "data.csv"))
```
### Descriptive summary

The mean number of caterpillars, with standard deviation, is summarised by
habitat.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Summarise the mean number of caterpillars with error
summary_data <- data %>%
  group_by(habitat) %>%
  summarise(
    mean_count = mean(caterpillar_count),
    sd = sd(caterpillar_count),
  )
```
### Removing erroneous records

Values from habitat D are removed, as these are known to be an error.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Remove values from habitat D as these are an error
filtered_data <- data %>%
  filter(habitat != "D")
```
## Statistical analysis

A Poisson generalised linear model is used to analyse caterpillar abundance
by habitat.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Run a Poisson general linear model to analyse caterpillar abundance varying with habitat. Produces numeric results in “Caterpillar Abundance”
model1 <- glm(caterpillar_count ~ habitat,
              family = poisson, data = filtered_data)

summary(model1)
```
Habitat significantly alters caterpillar abundance.

## Visualisation
```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Create figure 1, caterpillar count against habitat
figure1 <- ggplot(
  filtered_data,
  aes(x = habitat, y = caterpillar_count)
) +
  geom_boxplot() +
  theme_bw()

print(figure1)
```

Habitat E has the highest caterpillar count.

