---
title: "Caterpillar Abundance Analysis"
author: "Edward R. Ivimey-Cook"
output: html_document
---
  
```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Load packages  
library(dplyr)
library(ggplot2)
library(here)
```
## Load caterpillar abundance data (w/o local file paths)
```{r, echo=TRUE, message=FALSE, warning=FALSE}
data <- read.csv(here("caterpillar_data", "data.csv"))
```

## Summarise the mean number of caterpillars with error
```{r, echo=TRUE, message=FALSE, warning=FALSE}
summary_data <- data %>%
  group_by(habitat) %>%
  summarise(
    mean_count = mean(caterpillar_count),
    sd = sd(caterpillar_count),
  )
```

## Remove values from habitat D as these are an error
```{r, echo=TRUE, message=FALSE, warning=FALSE}
filtered_data <- data %>%
  filter(habitat != "D")
```

## Run a Poisson general linear model to analyse caterpillar abundance varying with habitat. Produces numeric results in “Caterpillar Abundance”
```{r, echo=TRUE, message=FALSE, warning=FALSE}
model1 <- glm(caterpillar_count ~ habitat,
              family = poisson, data = filtered_data)

summary(model1)
```
## Habitat significantly alters caterpillar abunndace

## Create figure 1, caterpillar count against habitat
```{r, echo=TRUE, message=FALSE, warning=FALSE}
figure1 <- ggplot(
  filtered_data,
  aes(x = habitat, y = caterpillar_count)
) +
  geom_boxplot() +
  theme_bw()

print(figure1)
```
## Habitat E has the highest caterpillar count


